---
title: "PERSUADE OUTPUT"
output:
  pdf_document:
    toc: yes
    keep_tex: TRUE
urlcolor: blue
---

```{r setup, include=FALSE}
#code is formatted using formatR::tidy_source(width.cutoff = 80)
knitr::opts_chunk$set(echo = TRUE, fig.align = "left", fig.path = "images/", dev = c('pdf', 'png'), dpi=300)
```

```{r preparation, echo=FALSE, warning = FALSE, message = FALSE, results = "hide"}
packages <- c("rms", "survival", "flexsurv", "muhaz", "survminer", "ggplot2", "data.table",
              "summarytools", "knitr", "kableExtra", "sft")
new.packages <- packages[!(packages %in% installed.packages()[, "Package"])]  #check for new packages
if (length(new.packages)) install.packages(new.packages)  #install new packages (if needed)
suppressPackageStartupMessages(lapply(packages, require, character.only = TRUE))  #load packages

# COLOUR PALETTE
n <- 8  #number of different colors (to be used for palette)
palette(rainbow(n = n, s = 1, v = 1, start = 0, end = max(1, n - 1)/n, alpha = 1))
```

```{r input, echo=FALSE, warning = FALSE, message = FALSE, results = "hide"}
# INPUT DATA
load("PERSUADE.RData")
```

\  
\  
Name (date): `r PERSUADE$name` (`r Sys.Date()`)\
\
[Link to PERSUADE GitHub page](https://github.com/Bram-R/PERSUADE)
\newpage

#Kaplan-Meier
```{r plot_KM, echo=FALSE, fig.height=8, fig.width=7}
evalq({
    suppressWarnings(if (ngroups > 1) {
        ggsurvplot(surv_fit(form, data = data.frame(years, status, group)), linetype = as.integer(c(1, "2222", "5212")), 
            legend.labs = group_names, risk.table = TRUE, conf.int = TRUE, surv.median.line = "hv", 
            ncensor.plot = TRUE, censor = FALSE, ggtheme = theme_light(), color = "strata", size = 0.5)
    } else {
        ggsurvplot(surv_fit(form, data = data.frame(years, status)), legend.labs = group_names, 
            risk.table = TRUE, conf.int = TRUE, surv.median.line = "hv", ncensor.plot = FALSE, 
            censor = TRUE, ggtheme = theme_light(), color = "black", size = 0.5)
    })
}, envir = PERSUADE)
```

\newpage

#Stratified models?
Should stratified parametric survival models be used?
```{r PH_assumption, echo=FALSE, eval=if (PERSUADE$ngroups>1) {TRUE} else {FALSE}, out.height = "29%", fig.width=11}
evalq({
    plot(cbind(log(km$time), log(-log(km$surv))), main = "A: LN(cumulative hazard)", 
        xlim = c(floor(min(log(km$time)[km$surv!=1])), ceiling(max(log(km$time)))),
        cex = 0.6, xlab = "LN(time)", ylab = "LN(cumulative hazard)", pch = c(1, 
            8, 9)[km_names], col = c("black", "lightgrey", "darkgrey")[km_names])
    
    plot(cox.zph(cox_reg, terms = FALSE)[1], col = "5", lwd = 2, xlab = "time", main = "B: Scaled Schoenfeld residuals", 
        ylab = "Beta(t) for group 2 vs group 1")
    abline(0, 0, lty = 3)
    abline(lm(cox.zph(cox_reg, terms = FALSE)$y[, 1] ~ cox.zph(cox_reg)$x)$coefficients, 
        col = "7", lty = 1, lwd = 2)
    legend(0, -2, col = c("5", "7"), lty = 1, cex = 0.8, legend = c("smoothed line (natural spline with df = 4)", 
        "regression line"))
    if (ngroups > 2) {
        plot(cox.zph(cox_reg, terms = FALSE)[2], col = "5", lwd = 2, xlab = "time", 
            main = "C: Scaled Schoenfeld residuals", ylab = "Beta(t) for group 3 vs group 1")
        abline(0, 0, lty = 3)
        abline(lm(cox.zph(cox_reg, terms = FALSE)$y[, 2] ~ cox.zph(cox_reg)$x)$coefficients, 
            col = "7", lty = 1, lwd = 2)
        legend(0, -2, col = c("5", "7"), lty = 1, cex = 0.8, legend = c("smoothed line (natural spline with df = 4)", 
            "regression line"))
    }
}, envir = PERSUADE)
```

\newpage

#Monotonic hazard models?
Should parametric survival models assuming a monotonic hazard rate (i.e. exponential, Weibull, Gompertz) be used?
```{r plot_hr, echo=FALSE}
evalq({
    plot(cbind(hr_smooth1$est.grid, hr_smooth1$haz.est), type = "l", col = "red", 
        lty = 1, lwd = 2, xlab = "time", ylab = "smoothed hazard rate", xlim = c(0, hr_max$time), 
        ylim = c(0, hr_max$hr_smooth))
    if (ngroups > 1) {
        lines(cbind(hr_smooth2$est.grid, hr_smooth2$haz.est), col = "green", lty = "2222", 
            lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(hr_smooth3$est.grid, hr_smooth3$haz.est), col = "blue", lty = "5212", 
            lwd = 2)
    }
    legend("topleft", legend = group_names, col = c("red", "green", "blue"), lty = as.integer(c(1, 
        "2222", "5212")), cex = 0.8)
}, envir = PERSUADE)
```

\newpage

#Standard parametric models?
Do standard parametric models provide an appropriate fit to the data?
```{r plot_parametric, echo=FALSE}
evalq({
    pie(rep(1, n), labels = lbls, col = rainbow(n = n, s = 1, v = 1, start = 0, end = max(1, 
        n - 1)/n, alpha = 1))
    kable_styling(kable(IC[order(IC$AIC), ], format = "latex", row.names = FALSE, 
        booktabs = TRUE, linesep = ""), latex_options = "striped")
}, envir = PERSUADE)
```

##Exponential
```{r expo, echo=FALSE, out.height = "30%", fig.width=11}
evalq({
    # constant hazard
    plot(km, lwd = 2, main = "A: Kaplan-Meier (Exponential)", xlab = "time", ylab = "survival", 
        col = c("black", "lightgrey", "darkgrey"))
    lines(cbind(time_pred, expo_pred[, 2]), col = "1", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(time_pred, expo_pred[, 3]), col = "1", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(time_pred, expo_pred[, 4]), col = "1", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km$time[km_names==1], rev(km$time[km_names==1])), 
            c(km$lower[km_names==1], rev(km$upper[km_names==1])), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km$time[km_names==2], rev(km$time[km_names==2])), 
              c(km$lower[km_names==2], rev(km$upper[km_names==2])), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km$time[km_names==3], rev(km$time[km_names==3])), 
              c(km$lower[km_names==3], rev(km$upper[km_names==3])), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    legend("bottomleft", legend = group_names, col = "1", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
    
    # annual transition probabilities
    plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), lwd = 1, lty = 1, type = "b", 
         cex = 0.6, pch = 1, col = "black", main = "B: Annual transition probability (Exponential)", 
         xlab = "time", ylab = "annual transition probability (smoothed)", 
         ylim = c(0, ceiling(km_tp$maxtp_smooth*10)/10), xlim = c(0, ceiling(km$maxtime*10)/10))
    lines(cbind(time_pred[-1], tp_gr_1[, 2]), col = "1", lty = 1, lwd = 2)
    if (ngroups > 1) {
      lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), col = "lightgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 8)
      lines(cbind(time_pred[-1], tp_gr_2[, 2]), col = "1", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
      lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), col = "darkgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 9)
      lines(cbind(time_pred[-1], tp_gr_3[, 2]), col = "1", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km_tp_gr_1$time, rev(km_tp_gr_1$time)), 
            c(km_tp_gr_1$tp_smooth_lower, rev(km_tp_gr_1$tp_smooth_upper)), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km_tp_gr_2$time, rev(km_tp_gr_2$time)), 
              c(km_tp_gr_2$tp_smooth_lower, rev(km_tp_gr_2$tp_smooth_upper)), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km_tp_gr_3$time, rev(km_tp_gr_3$time)), 
              c(km_tp_gr_3$tp_smooth_lower, rev(km_tp_gr_3$tp_smooth_upper)), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    legend("topleft", legend = group_names, col = "1", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)
    
    # diagnostics
    plot(cbind(km$time, -log(km$surv)), cex = 0.6, pch = c(1, 8, 9)[km_names], col = c("black", 
        "lightgrey", "darkgrey")[km_names], main = "C: Diagnostic plot (Exponential)", 
        xlab = "time", ylab = "cumulative hazard")
    lines(cbind(time_pred, -log(expo_pred[, 2])), col = "1", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(time_pred, -log(expo_pred[, 3])), col = "1", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(time_pred, -log(expo_pred[, 4])), col = "1", lty = "5212", lwd = 2)
    }
    legend("topleft", legend = group_names, col = "1", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
}, envir = PERSUADE)
```

##Weibull
```{r weib, echo=FALSE, out.height = "30%", fig.width=11}
evalq({
    # monotonically increasing or decreasing hazard
    plot(km, lwd = 2, main = "A: Kaplan-Meier (Weibull)", xlab = "time", ylab = "survival", 
        col = c("black", "lightgrey", "darkgrey"))
    lines(cbind(time_pred, weib_pred[, 2]), col = "2", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(time_pred, weib_pred[, 3]), col = "2", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(time_pred, weib_pred[, 4]), col = "2", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km$time[km_names==1], rev(km$time[km_names==1])), 
            c(km$lower[km_names==1], rev(km$upper[km_names==1])), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km$time[km_names==2], rev(km$time[km_names==2])), 
              c(km$lower[km_names==2], rev(km$upper[km_names==2])), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km$time[km_names==3], rev(km$time[km_names==3])), 
              c(km$lower[km_names==3], rev(km$upper[km_names==3])), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    legend("bottomleft", legend = group_names, col = "2", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
    
    # annual transition probabilities
    plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), lwd = 1, lty = 1, type = "b", 
        cex = 0.6, pch = 1, col = "black", main = "B: Annual transition probability (Weibull)", 
        xlab = "time", ylab = "annual transition probability (smoothed)", 
        ylim = c(0, ceiling(km_tp$maxtp_smooth*10)/10), xlim = c(0, ceiling(km$maxtime*10)/10))
    lines(cbind(time_pred[-1], tp_gr_1[, 3]), col = "2", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), col = "lightgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 8)
        lines(cbind(time_pred[-1], tp_gr_2[, 3]), col = "2", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), col = "darkgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 9)
        lines(cbind(time_pred[-1], tp_gr_3[, 3]), col = "2", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km_tp_gr_1$time, rev(km_tp_gr_1$time)), 
            c(km_tp_gr_1$tp_smooth_lower, rev(km_tp_gr_1$tp_smooth_upper)), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km_tp_gr_2$time, rev(km_tp_gr_2$time)), 
              c(km_tp_gr_2$tp_smooth_lower, rev(km_tp_gr_2$tp_smooth_upper)), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km_tp_gr_3$time, rev(km_tp_gr_3$time)), 
              c(km_tp_gr_3$tp_smooth_lower, rev(km_tp_gr_3$tp_smooth_upper)), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    legend("topleft", legend = group_names, col = "2", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
    
    # diagnostics
    plot(cbind(log(km$time), log(-log(km$surv))), cex = 0.6, pch = c(1, 8, 9)[km_names], 
        xlim = c(floor(min(log(km$time)[km$surv!=1])), ceiling(max(log(km$time)))),
        col = c("black", "lightgrey", "darkgrey")[km_names], main = "C: Diagnostic plot (Weibull)", 
        xlab = "LN(time)", ylab = "LN(cumulative hazard)")
    lines(cbind(log(time_pred), log(-log(weib_pred[, 2]))), col = "2", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(log(time_pred), log(-log(weib_pred[, 3]))), col = "2", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(log(time_pred), log(-log(weib_pred[, 4]))), col = "2", lty = "5212", lwd = 2)
    }
    legend("topleft", legend = group_names, col = "2", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
}, envir = PERSUADE)
```

##Gompertz
```{r gom, echo=FALSE, out.height = "30%", fig.width=11}
evalq({
    # monotonically increasing or decreasing hazard
    plot(km, lwd = 2, main = "A: Kaplan-Meier (Gompertz)", xlab = "time", ylab = "survival", 
        col = c("black", "lightgrey", "darkgrey"))
    lines(cbind(time_pred, gom_pred[, 2]), col = "3", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(time_pred, gom_pred[, 3]), col = "3", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(time_pred, gom_pred[, 4]), col = "3", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km$time[km_names==1], rev(km$time[km_names==1])), 
            c(km$lower[km_names==1], rev(km$upper[km_names==1])), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km$time[km_names==2], rev(km$time[km_names==2])), 
              c(km$lower[km_names==2], rev(km$upper[km_names==2])), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km$time[km_names==3], rev(km$time[km_names==3])), 
              c(km$lower[km_names==3], rev(km$upper[km_names==3])), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    legend("bottomleft", legend = group_names, col = "3", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
    
    # annual transition probabilities
    plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), lwd = 1, lty = 1, type = "b", 
         cex = 0.6, pch = 1, col = "black", main = "B: Annual transition probability (Gompertz)", 
         xlab = "time", ylab = "annual transition probability (smoothed)",
         ylim = c(0, ceiling(km_tp$maxtp_smooth*10)/10), xlim = c(0, ceiling(km$maxtime*10)/10))
    lines(cbind(time_pred[-1], tp_gr_1[, 4]), col = "3", lty = 1, lwd = 2)
    if (ngroups > 1) {
      lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), col = "lightgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 8)
      lines(cbind(time_pred[-1], tp_gr_2[, 4]), col = "3", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
      lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), col = "darkgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 9)
      lines(cbind(time_pred[-1], tp_gr_3[, 4]), col = "3", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km_tp_gr_1$time, rev(km_tp_gr_1$time)), 
            c(km_tp_gr_1$tp_smooth_lower, rev(km_tp_gr_1$tp_smooth_upper)), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km_tp_gr_2$time, rev(km_tp_gr_2$time)), 
              c(km_tp_gr_2$tp_smooth_lower, rev(km_tp_gr_2$tp_smooth_upper)), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km_tp_gr_3$time, rev(km_tp_gr_3$time)), 
              c(km_tp_gr_3$tp_smooth_lower, rev(km_tp_gr_3$tp_smooth_upper)), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    legend("topleft", legend = group_names, col = "3", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)
    
    # diagnostics
    plot(cbind(c(hr_smooth1$est.grid, if (ngroups > 1) {
        hr_smooth2$est.grid
    } else {
        NA
    }, if (ngroups > 2) {
        hr_smooth3$est.grid
    } else {
        NA
    }), c(log(hr_smooth1$haz.est), if (ngroups > 1) {
        log(hr_smooth2$haz.est)
    } else {
        NA
    }, if (ngroups > 2) {
        log(hr_smooth3$haz.est)
    } else {
        NA
    })), cex = 0.6, main = "C: Diagnostic plot (Gompertz)", xlab = "time", ylab = "LN(smoothed hazard)", 
        pch = c(1, 8, 9)[hr_names], col = c("black", "lightgrey", "darkgrey")[hr_names])
    lines(cbind(gom_est_h[[1]]$time, log(gom_est_h[[1]]$est)), col = "3", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(gom_est_h[[2]]$time, log(gom_est_h[[2]]$est)), col = "3", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(gom_est_h[[3]]$time, log(gom_est_h[[3]]$est)), col = "3", lty = "5212", lwd = 2)
    }
    legend("topleft", legend = group_names, col = "3", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
}, envir = PERSUADE)
```

##Log-normal
```{r lnorm, echo=FALSE, out.height = "30%", fig.width=11}
evalq({
    # hazard can be non-monotonic with respect to time; initially an increasing
    # hazard, followed by a decreasing hazard
    plot(km, lwd = 2, main = "A: Kaplan-Meier (Log-normal)", xlab = "time", ylab = "survival", 
        col = c("black", "lightgrey", "darkgrey"))
    lines(cbind(time_pred, lnorm_pred[, 2]), col = "4", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(time_pred, lnorm_pred[, 3]), col = "4", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(time_pred, lnorm_pred[, 4]), col = "4", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km$time[km_names==1], rev(km$time[km_names==1])), 
            c(km$lower[km_names==1], rev(km$upper[km_names==1])), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km$time[km_names==2], rev(km$time[km_names==2])), 
              c(km$lower[km_names==2], rev(km$upper[km_names==2])), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km$time[km_names==3], rev(km$time[km_names==3])), 
              c(km$lower[km_names==3], rev(km$upper[km_names==3])), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    legend("bottomleft", legend = group_names, col = "4", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
    
    # annual transition probabilities
    plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), lwd = 1, lty = 1, type = "b", 
         cex = 0.6, pch = 1, col = "black", main = "B: Annual transition probability (Log-normal)", 
         xlab = "time", ylab = "annual transition probability (smoothed)", 
         ylim = c(0, ceiling(km_tp$maxtp_smooth*10)/10), xlim = c(0, ceiling(km$maxtime*10)/10))
    lines(cbind(time_pred[-1], tp_gr_1[, 5]), col = "4", lty = 1, lwd = 2)
    if (ngroups > 1) {
      lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), col = "lightgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 8)
      lines(cbind(time_pred[-1], tp_gr_2[, 5]), col = "4", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
      lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), col = "darkgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 9)
      lines(cbind(time_pred[-1], tp_gr_3[, 5]), col = "4", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km_tp_gr_1$time, rev(km_tp_gr_1$time)), 
            c(km_tp_gr_1$tp_smooth_lower, rev(km_tp_gr_1$tp_smooth_upper)), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km_tp_gr_2$time, rev(km_tp_gr_2$time)), 
              c(km_tp_gr_2$tp_smooth_lower, rev(km_tp_gr_2$tp_smooth_upper)), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km_tp_gr_3$time, rev(km_tp_gr_3$time)), 
              c(km_tp_gr_3$tp_smooth_lower, rev(km_tp_gr_3$tp_smooth_upper)), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    legend("topleft", legend = group_names, col = "4", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)
    
    # diagnostics
    plot(cbind(log(km$time), -qnorm(km$surv, mean = 0, sd = 1, lower.tail = TRUE, 
        log.p = FALSE)), cex = 0.6, main = "C: Diagnostic plot (Log-normal)", xlab = "LN(time)",
        xlim = c(floor(min(log(km$time)[km$surv!=1])), ceiling(max(log(km$time)))),
        pch = c(1, 8, 9)[km_names], col = c("black", "lightgrey", "darkgrey")[km_names], 
        ylab = "- standard normal quartiles")
    lines(cbind(log(time_pred), -qnorm(lnorm_pred[, 2], mean = 0, sd = 1, lower.tail = TRUE, 
        log.p = FALSE)), col = "4", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(log(time_pred), -qnorm(lnorm_pred[, 3], mean = 0, sd = 1, lower.tail = TRUE, 
            log.p = FALSE)), col = "4", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(log(time_pred), -qnorm(lnorm_pred[, 4], mean = 0, sd = 1, lower.tail = TRUE, 
            log.p = FALSE)), col = "4", lty = "5212", lwd = 2)
    }
    legend("topleft", legend = group_names, col = "4", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
}, envir = PERSUADE)
```

##Log-logistic
```{r llog, echo=FALSE, out.height = "30%", fig.width=11}
evalq({
    # hazard can be non-monotonic with respect to time; initially an increasing
    # hazard, followed by a decreasing hazard
    plot(km, lwd = 2, main = "A: Kaplan-Meier (Log-logistic)", xlab = "time", ylab = "survival", 
        col = c("black", "lightgrey", "darkgrey"))
    lines(cbind(time_pred, llog_pred[, 2]), col = "5", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(time_pred, llog_pred[, 3]), col = "5", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(time_pred, llog_pred[, 4]), col = "5", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km$time[km_names==1], rev(km$time[km_names==1])), 
            c(km$lower[km_names==1], rev(km$upper[km_names==1])), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km$time[km_names==2], rev(km$time[km_names==2])), 
              c(km$lower[km_names==2], rev(km$upper[km_names==2])), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km$time[km_names==3], rev(km$time[km_names==3])), 
              c(km$lower[km_names==3], rev(km$upper[km_names==3])), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    legend("bottomleft", legend = group_names, col = "5", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
    
    # annual transition probabilities
    plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), lwd = 1, lty = 1, type = "b", 
         cex = 0.6, pch = 1, col = "black", main = "B: Annual transition probability (Log-logistic)", 
         xlab = "time", ylab = "annual transition probability (smoothed)", 
         ylim = c(0, ceiling(km_tp$maxtp_smooth*10)/10), xlim = c(0, ceiling(km$maxtime*10)/10))
    lines(cbind(time_pred[-1], tp_gr_1[, 6]), col = "5", lty = 1, lwd = 2)
    if (ngroups > 1) {
      lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), col = "lightgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 8)
      lines(cbind(time_pred[-1], tp_gr_2[, 6]), col = "5", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
      lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), col = "darkgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 9)
      lines(cbind(time_pred[-1], tp_gr_3[, 6]), col = "5", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km_tp_gr_1$time, rev(km_tp_gr_1$time)), 
            c(km_tp_gr_1$tp_smooth_lower, rev(km_tp_gr_1$tp_smooth_upper)), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km_tp_gr_2$time, rev(km_tp_gr_2$time)), 
              c(km_tp_gr_2$tp_smooth_lower, rev(km_tp_gr_2$tp_smooth_upper)), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km_tp_gr_3$time, rev(km_tp_gr_3$time)), 
              c(km_tp_gr_3$tp_smooth_lower, rev(km_tp_gr_3$tp_smooth_upper)), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    legend("topleft", legend = group_names, col = "5", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)
    
    # diagnostics
    plot(cbind(log(km$time), -log(km$surv/(1 - km$surv))), cex = 0.6, pch = c(1, 
        8, 9)[km_names], col = c("black", "lightgrey", "darkgrey")[km_names], 
        xlim = c(floor(min(log(km$time)[km$surv!=1])), ceiling(max(log(km$time)))),
        main = "C: Diagnostic plot (Log-logistic)", xlab = "LN(time)", ylab = "-LN(survival odds)")
    lines(cbind(log(time_pred), -log(llog_pred[, 2]/(1 - llog_pred[, 2]))), col = "5", 
        lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(log(time_pred), -log(llog_pred[, 3]/(1 - llog_pred[, 3]))), col = "5", 
            lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(log(time_pred), -log(llog_pred[, 4]/(1 - llog_pred[, 4]))), col = "5", 
            lty = "5212", lwd = 2)
    }
    legend("topleft", legend = group_names, col = "5", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
}, envir = PERSUADE)
```

##Gamma
```{r gam, echo=FALSE, out.height = "30%", fig.width=11}
evalq({
    # hazard rate has (inverted) 'U' shape over time
    plot(km, lwd = 2, main = "A: Kaplan-Meier (Gamma)", xlab = "time", ylab = "survival", 
        col = c("black", "lightgrey", "darkgrey"))
    lines(cbind(time_pred, gam_pred[, 2]), col = "6", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(time_pred, gam_pred[, 3]), col = "6", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(time_pred, gam_pred[, 4]), col = "6", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km$time[km_names==1], rev(km$time[km_names==1])), 
            c(km$lower[km_names==1], rev(km$upper[km_names==1])), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km$time[km_names==2], rev(km$time[km_names==2])), 
              c(km$lower[km_names==2], rev(km$upper[km_names==2])), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km$time[km_names==3], rev(km$time[km_names==3])), 
              c(km$lower[km_names==3], rev(km$upper[km_names==3])), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    legend("bottomleft", legend = group_names, col = "6", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
    
    # annual transition probabilities
    plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), lwd = 1, lty = 1, type = "b", 
         cex = 0.6, pch = 1, col = "black", main = "B: Annual transition probability (Gamma)", 
         xlab = "time", ylab = "annual transition probability (smoothed)", 
         ylim = c(0, ceiling(km_tp$maxtp_smooth*10)/10), xlim = c(0, ceiling(km$maxtime*10)/10))
    lines(cbind(time_pred[-1], tp_gr_1[, 7]), col = "6", lty = 1, lwd = 2)
    if (ngroups > 1) {
      lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), col = "lightgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 8)
      lines(cbind(time_pred[-1], tp_gr_2[, 7]), col = "6", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
      lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), col = "darkgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 9)
      lines(cbind(time_pred[-1], tp_gr_3[, 7]), col = "6", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km_tp_gr_1$time, rev(km_tp_gr_1$time)), 
            c(km_tp_gr_1$tp_smooth_lower, rev(km_tp_gr_1$tp_smooth_upper)), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km_tp_gr_2$time, rev(km_tp_gr_2$time)), 
              c(km_tp_gr_2$tp_smooth_lower, rev(km_tp_gr_2$tp_smooth_upper)), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km_tp_gr_3$time, rev(km_tp_gr_3$time)), 
              c(km_tp_gr_3$tp_smooth_lower, rev(km_tp_gr_3$tp_smooth_upper)), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    legend("topleft", legend = group_names, col = "6", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)
    
    # diagnostics --> not linear!
    plot(cbind(log(km$time), log(-log(km$surv))), cex = 0.6, pch = c(1, 8, 9)[km_names],
        xlim = c(floor(min(log(km$time)[km$surv!=1])), ceiling(max(log(km$time)))),
        col = c("black", "lightgrey", "darkgrey")[km_names], main = "C: Diagnostic plot (Gamma)", 
        xlab = "LN(time)", ylab = "LN(cumulative hazard)")
    lines(cbind(log(time_pred), log(-log(gam_pred[, 2]))), col = "6", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(log(time_pred), log(-log(gam_pred[, 3]))), col = "6", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(log(time_pred), log(-log(gam_pred[, 4]))), col = "6", lty = "5212", lwd = 2)
    }
    legend("topleft", legend = group_names, col = "6", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
}, envir = PERSUADE)
```

##Generalised Gamma 
```{r ggam, echo=FALSE, out.height = "30%", fig.width=11}
evalq({
    # flexible: can take form of Exponential, Weibull, Gamma or Log-normal
    plot(km, lwd = 2, main = "A: Kaplan-Meier (Generalised gamma)", xlab = "time", 
        ylab = "survival", col = c("black", "lightgrey", "darkgrey"))
    lines(cbind(time_pred, ggam_pred[, 2]), col = "7", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(time_pred, ggam_pred[, 3]), col = "7", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(time_pred, ggam_pred[, 4]), col = "7", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km$time[km_names==1], rev(km$time[km_names==1])), 
            c(km$lower[km_names==1], rev(km$upper[km_names==1])), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km$time[km_names==2], rev(km$time[km_names==2])), 
              c(km$lower[km_names==2], rev(km$upper[km_names==2])), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km$time[km_names==3], rev(km$time[km_names==3])), 
              c(km$lower[km_names==3], rev(km$upper[km_names==3])), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    legend("bottomleft", legend = group_names, col = "7", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
    
    # annual transition probabilities
    plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), lwd = 1, lty = 1, type = "b", 
         cex = 0.6, pch = 1, col = "black", main = "B: Annual transition probability (Generalised Gamma)", 
         xlab = "time", ylab = "annual transition probability (smoothed)", 
         ylim = c(0, ceiling(km_tp$maxtp_smooth*10)/10), xlim = c(0, ceiling(km$maxtime*10)/10))
    lines(cbind(time_pred[-1], tp_gr_1[, 8]), col = "7", lty = 1, lwd = 2)
    if (ngroups > 1) {
      lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), col = "lightgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 8)
      lines(cbind(time_pred[-1], tp_gr_2[, 8]), col = "7", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
      lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), col = "darkgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 9)
      lines(cbind(time_pred[-1], tp_gr_3[, 8]), col = "7", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km_tp_gr_1$time, rev(km_tp_gr_1$time)), 
            c(km_tp_gr_1$tp_smooth_lower, rev(km_tp_gr_1$tp_smooth_upper)), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km_tp_gr_2$time, rev(km_tp_gr_2$time)), 
              c(km_tp_gr_2$tp_smooth_lower, rev(km_tp_gr_2$tp_smooth_upper)), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km_tp_gr_3$time, rev(km_tp_gr_3$time)), 
              c(km_tp_gr_3$tp_smooth_lower, rev(km_tp_gr_3$tp_smooth_upper)), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    legend("topleft", legend = group_names, col = "7", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)
    
    # diagnostics --> not linear!
    plot(cbind(log(km$time), log(-log(km$surv))), cex = 0.6, pch = c(1, 8, 9)[km_names],
        xlim = c(floor(min(log(km$time)[km$surv!=1])), ceiling(max(log(km$time)))),
        col = c("black", "lightgrey", "darkgrey")[km_names], main = "C: Diagnostic plot (Generalised gamma)", 
        xlab = "LN(time)", ylab = "LN(cumulative hazard)")
    lines(cbind(log(time_pred), log(-log(ggam_pred[, 2]))), col = "7", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(log(time_pred), log(-log(ggam_pred[, 3]))), col = "7", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(log(time_pred), log(-log(ggam_pred[, 4]))), col = "7", lty = "5212", lwd = 2)
    }
    legend("topleft", legend = group_names, col = "7", lty = as.integer(c(1, "2222", 
        "5212")), cex = 0.8)
}, envir = PERSUADE)
```

\newpage

#Parametric spline models?
If standard parametric models are not appropriate, are spline models a more appropriate fit to the data?

```{r spline, echo=FALSE, eval = PERSUADE$show_spline}
evalq({
    # k=0 gives a Weibull, log-logistic or log-normal model, if 'scale' is 'hazard',
    # 'odds' or 'normal' respectively. The knots are then chosen as equally-spaced
    # quantiles of the log uncensored survival times, for example, at the median with
    # one knot, or at the 33% and 67% quantiles of log time (or time, see
    # 'timescale') with two knots.
    
    pie(rep(1, n), labels = lbls_spline, col = rainbow(n = n, s = 1, v = 1, start = 0, 
        end = max(1, n - 1)/n, alpha = 1))
    
    kable_styling(kable(rbind(IC, IC_spl)[order(rbind(IC, IC_spl)$AIC), ], format = "latex", 
        row.names = FALSE, booktabs = TRUE, linesep = ""), latex_options = "striped")
}, envir = PERSUADE)
```

##Spline hazard 1 knot
```{r spline_hazard1, echo=FALSE, out.height = "30%", fig.width=11, eval = PERSUADE$show_spline}
evalq({
    # 1 knot
    plot(km, lwd = 2, main = "A: Kaplan-Meier (Spline, 1 knot, hazard scale)", xlab = "time", 
        ylab = "survival", col = c("black", "lightgrey", "darkgrey"))
    lines(cbind(time_pred, spl_hazard1_pred[, 2]), col = "1", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(time_pred, spl_hazard1_pred[, 3]), col = "1", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(time_pred, spl_hazard1_pred[, 4]), col = "1", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km$time[km_names==1], rev(km$time[km_names==1])), 
            c(km$lower[km_names==1], rev(km$upper[km_names==1])), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km$time[km_names==2], rev(km$time[km_names==2])), 
              c(km$lower[km_names==2], rev(km$upper[km_names==2])), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km$time[km_names==3], rev(km$time[km_names==3])), 
              c(km$lower[km_names==3], rev(km$upper[km_names==3])), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    abline(v = exp(spl_hazard1$knots[2]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_hazard1$knots[1]), lty = 3, lwd = 1)
    abline(v = exp(spl_hazard1$knots[3]), lty = 3, lwd = 1)
    legend("bottomleft", legend = group_names, col = "1", lty = as.integer(c(1, "2222", 
                                                                             "5212")), cex = 0.8)
    
    # annual transition probabilities
    plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), lwd = 1, lty = 1, type = "b", 
         cex = 0.6, pch = 1, col = "black", main = "B: Annual transition probability (Spline, 1 knot, hazard scale)", 
         xlab = "time", ylab = "annual transition probability (smoothed)", 
         ylim = c(0, ceiling(km_tp$maxtp_smooth*10)/10), xlim = c(0, ceiling(km$maxtime*10)/10))
    lines(cbind(time_pred[-1], tp_gr_1[, 9]), col = "1", lty = 1, lwd = 2)
    if (ngroups > 1) {
      lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), col = "lightgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 8)
      lines(cbind(time_pred[-1], tp_gr_2[, 9]), col = "1", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
      lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), col = "darkgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 9)
      lines(cbind(time_pred[-1], tp_gr_3[, 9]), col = "1", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km_tp_gr_1$time, rev(km_tp_gr_1$time)), 
            c(km_tp_gr_1$tp_smooth_lower, rev(km_tp_gr_1$tp_smooth_upper)), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km_tp_gr_2$time, rev(km_tp_gr_2$time)), 
              c(km_tp_gr_2$tp_smooth_lower, rev(km_tp_gr_2$tp_smooth_upper)), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km_tp_gr_3$time, rev(km_tp_gr_3$time)), 
              c(km_tp_gr_3$tp_smooth_lower, rev(km_tp_gr_3$tp_smooth_upper)), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    abline(v = exp(spl_hazard1$knots[2]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_hazard1$knots[1]), lty = 3, lwd = 1)
    abline(v = exp(spl_hazard1$knots[3]), lty = 3, lwd = 1)
    legend("topleft", legend = group_names, col = "1", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)

    # diagnostics
    plot(cbind(log(km$time), log(-log(km$surv))), cex = 0.6, pch = c(1, 8, 9)[km_names],
        xlim = c(floor(min(log(km$time)[km$surv!=1])), ceiling(max(log(km$time)))),
        col = c("black", "lightgrey", "darkgrey")[km_names], main = "C: Diagnostic plot (Spline, 1 knot, hazard scale)", 
        xlab = "LN(time)", ylab = "LN(cumulative hazard)")
    lines(cbind(log(time_pred), log(-log(spl_hazard1_pred[, 2]))), col = "1", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(log(time_pred), log(-log(spl_hazard1_pred[, 3]))), col = "1", 
            lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(log(time_pred), log(-log(spl_hazard1_pred[, 4]))), col = "1", 
            lty = "5212", lwd = 2)
    }
    abline(v = spl_hazard1$knots[2], lty = 2, lwd = 1.5)
    abline(v = spl_hazard1$knots[1], lty = 3, lwd = 1)
    abline(v = spl_hazard1$knots[3], lty = 3, lwd = 1)
    legend("topleft", legend = group_names, col = "1", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)
}, envir = PERSUADE)
```

##Spline hazard 2 knots
```{r spline_hazard2, echo=FALSE, out.height = "30%", fig.width=11, eval = PERSUADE$show_spline}
evalq({
    # 2 knots
    plot(km, lwd = 2, main = "A: Kaplan-Meier (Spline, 2 knots, hazard scale)", xlab = "time", 
        ylab = "survival", col = c("black", "lightgrey", "darkgrey"))
    lines(cbind(time_pred, spl_hazard2_pred[, 2]), col = "2", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(time_pred, spl_hazard2_pred[, 3]), col = "2", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(time_pred, spl_hazard2_pred[, 4]), col = "2", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km$time[km_names==1], rev(km$time[km_names==1])), 
            c(km$lower[km_names==1], rev(km$upper[km_names==1])), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km$time[km_names==2], rev(km$time[km_names==2])), 
              c(km$lower[km_names==2], rev(km$upper[km_names==2])), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km$time[km_names==3], rev(km$time[km_names==3])), 
              c(km$lower[km_names==3], rev(km$upper[km_names==3])), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    abline(v = exp(spl_hazard2$knots[2]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_hazard2$knots[3]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_hazard2$knots[1]), lty = 3, lwd = 1)
    abline(v = exp(spl_hazard2$knots[4]), lty = 3, lwd = 1)
    legend("bottomleft", legend = group_names, col = "2", lty = as.integer(c(1, "2222", 
                                                                             "5212")), cex = 0.8)
    
    # annual transition probabilities
    plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), lwd = 1, lty = 1, type = "b", 
         cex = 0.6, pch = 1, col = "black", main = "B: Annual transition probability (Spline, 2 knots, hazard scale)", 
         xlab = "time", ylab = "annual transition probability (smoothed)", 
         ylim = c(0, ceiling(km_tp$maxtp_smooth*10)/10), xlim = c(0, ceiling(km$maxtime*10)/10))
    lines(cbind(time_pred[-1], tp_gr_1[, 10]), col = "2", lty = 1, lwd = 2)
    if (ngroups > 1) {
      lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), col = "lightgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 8)
      lines(cbind(time_pred[-1], tp_gr_2[, 10]), col = "2", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
      lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), col = "darkgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 9)
      lines(cbind(time_pred[-1], tp_gr_3[, 10]), col = "2", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km_tp_gr_1$time, rev(km_tp_gr_1$time)), 
            c(km_tp_gr_1$tp_smooth_lower, rev(km_tp_gr_1$tp_smooth_upper)), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km_tp_gr_2$time, rev(km_tp_gr_2$time)), 
              c(km_tp_gr_2$tp_smooth_lower, rev(km_tp_gr_2$tp_smooth_upper)), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km_tp_gr_3$time, rev(km_tp_gr_3$time)), 
              c(km_tp_gr_3$tp_smooth_lower, rev(km_tp_gr_3$tp_smooth_upper)), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    abline(v = exp(spl_hazard2$knots[2]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_hazard2$knots[3]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_hazard2$knots[1]), lty = 3, lwd = 1)
    abline(v = exp(spl_hazard2$knots[4]), lty = 3, lwd = 1)
    legend("topleft", legend = group_names, col = "2", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)

    # diagnostics
    plot(cbind(log(km$time), log(-log(km$surv))), cex = 0.6, pch = c(1, 8, 9)[km_names],
        xlim = c(floor(min(log(km$time)[km$surv!=1])), ceiling(max(log(km$time)))),
        col = c("black", "lightgrey", "darkgrey")[km_names], main = "C: Diagnostic plot (Spline, 2 knots, hazard scale)", 
        xlab = "LN(time)", ylab = "LN(cumulative hazard)")
    lines(cbind(log(time_pred), log(-log(spl_hazard2_pred[, 2]))), col = "2", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(log(time_pred), log(-log(spl_hazard2_pred[, 3]))), col = "2", 
            lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(log(time_pred), log(-log(spl_hazard2_pred[, 4]))), col = "2", 
            lty = "5212", lwd = 2)
    }
    abline(v = spl_hazard2$knots[2], lty = 2, lwd = 1.5)
    abline(v = spl_hazard2$knots[3], lty = 2, lwd = 1.5)
    abline(v = spl_hazard2$knots[1], lty = 3, lwd = 1)
    abline(v = spl_hazard2$knots[4], lty = 3, lwd = 1)
    legend("topleft", legend = group_names, col = "2", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)
}, envir = PERSUADE)
```

##Spline odds 1 knot
```{r spline_odds1, echo=FALSE, out.height = "30%", fig.width=11, eval = PERSUADE$show_spline}
evalq({
    # 1 knot
    plot(km, lwd = 2, main = "A: Kaplan-Meier (Spline, 1 knot, odds scale)", xlab = "time", 
        ylab = "survival", col = c("black", "lightgrey", "darkgrey"))
    lines(cbind(time_pred, spl_odds1_pred[, 2]), col = "3", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(time_pred, spl_odds1_pred[, 3]), col = "3", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(time_pred, spl_odds1_pred[, 4]), col = "3", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km$time[km_names==1], rev(km$time[km_names==1])), 
            c(km$lower[km_names==1], rev(km$upper[km_names==1])), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km$time[km_names==2], rev(km$time[km_names==2])), 
              c(km$lower[km_names==2], rev(km$upper[km_names==2])), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km$time[km_names==3], rev(km$time[km_names==3])), 
              c(km$lower[km_names==3], rev(km$upper[km_names==3])), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    abline(v = exp(spl_odds1$knots[2]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_odds1$knots[1]), lty = 3, lwd = 1)
    abline(v = exp(spl_odds1$knots[3]), lty = 3, lwd = 1)
    legend("bottomleft", legend = group_names, col = "3", lty = as.integer(c(1, "2222", 
                                                                             "5212")), cex = 0.8)
    
    # annual transition probabilities
    plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), lwd = 1, lty = 1, type = "b", 
         cex = 0.6, pch = 1, col = "black", main = "B: Annual transition probability (Spline, 1 knot, odds scale)", 
         xlab = "time", ylab = "annual transition probability (smoothed)", 
         ylim = c(0, ceiling(km_tp$maxtp_smooth*10)/10), xlim = c(0, ceiling(km$maxtime*10)/10))
    lines(cbind(time_pred[-1], tp_gr_1[, 11]), col = "3", lty = 1, lwd = 2)
    if (ngroups > 1) {
      lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), col = "lightgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 8)
      lines(cbind(time_pred[-1], tp_gr_2[, 11]), col = "3", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
      lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), col = "darkgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 9)
      lines(cbind(time_pred[-1], tp_gr_3[, 11]), col = "3", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km_tp_gr_1$time, rev(km_tp_gr_1$time)), 
            c(km_tp_gr_1$tp_smooth_lower, rev(km_tp_gr_1$tp_smooth_upper)), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km_tp_gr_2$time, rev(km_tp_gr_2$time)), 
              c(km_tp_gr_2$tp_smooth_lower, rev(km_tp_gr_2$tp_smooth_upper)), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km_tp_gr_3$time, rev(km_tp_gr_3$time)), 
              c(km_tp_gr_3$tp_smooth_lower, rev(km_tp_gr_3$tp_smooth_upper)), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    abline(v = exp(spl_odds1$knots[2]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_odds1$knots[1]), lty = 3, lwd = 1)
    abline(v = exp(spl_odds1$knots[3]), lty = 3, lwd = 1)
    legend("topleft", legend = group_names, col = "3", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)

    # diagnostics
    plot(cbind(log(km$time), -log(km$surv/(1 - km$surv))), cex = 0.6, pch = c(1, 
        8, 9)[km_names], col = c("black", "lightgrey", "darkgrey")[km_names], 
        main = "C: Diagnostic plot (Spline, 1 knot, odds scale)", 
        xlim = c(floor(min(log(km$time)[km$surv!=1])), ceiling(max(log(km$time)))),
        xlab = "LN(time)", ylab = "-LN(survival odds)")
    lines(cbind(log(time_pred), -log(spl_odds1_pred[, 2]/(1 - spl_odds1_pred[, 2]))), 
        col = "3", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(log(time_pred), -log(spl_odds1_pred[, 3]/(1 - spl_odds1_pred[, 
            3]))), col = "3", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(log(time_pred), -log(spl_odds1_pred[, 4]/(1 - spl_odds1_pred[, 
            4]))), col = "3", lty = "5212", lwd = 2)
    }
    abline(v = spl_odds1$knots[2], lty = 2, lwd = 1.5)
    abline(v = spl_odds1$knots[1], lty = 3, lwd = 1)
    abline(v = spl_odds1$knots[3], lty = 3, lwd = 1)
    legend("topleft", legend = group_names, col = "3", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)
}, envir = PERSUADE)
```

##Spline odds 2 knots
```{r spline_odds2, echo=FALSE, out.height = "30%", fig.width=11, eval = PERSUADE$show_spline}
evalq({
    # 2 knots
    plot(km, lwd = 2, main = "A: Kaplan-Meier (Spline, 2 knots, odds scale)", xlab = "time", 
        ylab = "survival", col = c("black", "lightgrey", "darkgrey"))
    lines(cbind(time_pred, spl_odds2_pred[, 2]), col = "4", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(time_pred, spl_odds2_pred[, 3]), col = "4", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(time_pred, spl_odds2_pred[, 4]), col = "4", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km$time[km_names==1], rev(km$time[km_names==1])), 
            c(km$lower[km_names==1], rev(km$upper[km_names==1])), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km$time[km_names==2], rev(km$time[km_names==2])), 
              c(km$lower[km_names==2], rev(km$upper[km_names==2])), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km$time[km_names==3], rev(km$time[km_names==3])), 
              c(km$lower[km_names==3], rev(km$upper[km_names==3])), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    abline(v = exp(spl_odds2$knots[2]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_odds2$knots[3]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_odds2$knots[1]), lty = 3, lwd = 1)
    abline(v = exp(spl_odds2$knots[4]), lty = 3, lwd = 1)
    legend("bottomleft", legend = group_names, col = "4", lty = as.integer(c(1, "2222", 
                                                                             "5212")), cex = 0.8)
    
    # annual transition probabilities
    plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), lwd = 1, lty = 1, type = "b", 
         cex = 0.6, pch = 1, col = "black", main = "B: Annual transition probability (Spline, 2 knots, odds scale)", 
         xlab = "time", ylab = "annual transition probability (smoothed)", 
         ylim = c(0, ceiling(km_tp$maxtp_smooth*10)/10), xlim = c(0, ceiling(km$maxtime*10)/10))
    lines(cbind(time_pred[-1], tp_gr_1[, 12]), col = "4", lty = 1, lwd = 2)
    if (ngroups > 1) {
      lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), col = "lightgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 8)
      lines(cbind(time_pred[-1], tp_gr_2[, 12]), col = "4", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
      lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), col = "darkgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 9)
      lines(cbind(time_pred[-1], tp_gr_3[, 12]), col = "4", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km_tp_gr_1$time, rev(km_tp_gr_1$time)), 
            c(km_tp_gr_1$tp_smooth_lower, rev(km_tp_gr_1$tp_smooth_upper)), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km_tp_gr_2$time, rev(km_tp_gr_2$time)), 
              c(km_tp_gr_2$tp_smooth_lower, rev(km_tp_gr_2$tp_smooth_upper)), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km_tp_gr_3$time, rev(km_tp_gr_3$time)), 
              c(km_tp_gr_3$tp_smooth_lower, rev(km_tp_gr_3$tp_smooth_upper)), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    abline(v = exp(spl_odds2$knots[2]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_odds2$knots[3]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_odds2$knots[1]), lty = 3, lwd = 1)
    abline(v = exp(spl_odds2$knots[4]), lty = 3, lwd = 1)
    legend("topleft", legend = group_names, col = "4", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)


    # diagnostics
    plot(cbind(log(km$time), -log(km$surv/(1 - km$surv))), cex = 0.6, pch = c(1, 
        8, 9)[km_names], col = c("black", "lightgrey", "darkgrey")[km_names], 
        main = "C: Diagnostic plot (Spline, 2 knots, odds scale)", 
        xlim = c(floor(min(log(km$time)[km$surv!=1])), ceiling(max(log(km$time)))),
        xlab = "LN(time)", ylab = "-LN(survival odds)")
    lines(cbind(log(time_pred), -log(spl_odds2_pred[, 2]/(1 - spl_odds2_pred[, 2]))), 
        col = "4", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(log(time_pred), -log(spl_odds2_pred[, 3]/(1 - spl_odds2_pred[, 
            3]))), col = "4", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(log(time_pred), -log(spl_odds2_pred[, 4]/(1 - spl_odds2_pred[, 
            4]))), col = "4", lty = "5212", lwd = 2)
    }
    abline(v = spl_odds2$knots[2], lty = 2, lwd = 1.5)
    abline(v = spl_odds2$knots[3], lty = 2, lwd = 1.5)
    abline(v = spl_odds2$knots[1], lty = 3, lwd = 1)
    abline(v = spl_odds2$knots[4], lty = 3, lwd = 1)
    legend("topleft", legend = group_names, col = "4", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)
}, envir = PERSUADE)
```

##Spline normal 1 knot
```{r spline_norm1, echo=FALSE, out.height = "30%", fig.width=11, eval = PERSUADE$show_spline}
evalq({
    # 1 knot
    plot(km, lwd = 2, main = "A: Kaplan-Meier (Spline, 1 knot, normal scale)", xlab = "time", 
        ylab = "survival", col = c("black", "lightgrey", "darkgrey"))
    lines(cbind(time_pred, spl_normal1_pred[, 2]), col = "5", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(time_pred, spl_normal1_pred[, 3]), col = "5", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(time_pred, spl_normal1_pred[, 4]), col = "5", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km$time[km_names==1], rev(km$time[km_names==1])), 
            c(km$lower[km_names==1], rev(km$upper[km_names==1])), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km$time[km_names==2], rev(km$time[km_names==2])), 
              c(km$lower[km_names==2], rev(km$upper[km_names==2])), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km$time[km_names==3], rev(km$time[km_names==3])), 
              c(km$lower[km_names==3], rev(km$upper[km_names==3])), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    abline(v = exp(spl_normal1$knots[2]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_normal1$knots[1]), lty = 3, lwd = 1)
    abline(v = exp(spl_normal1$knots[3]), lty = 3, lwd = 1)
    legend("bottomleft", legend = group_names, col = "5", lty = as.integer(c(1, "2222", 
                                                                             "5212")), cex = 0.8)
    
    # annual transition probabilities
    plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), lwd = 1, lty = 1, type = "b", 
         cex = 0.6, pch = 1, col = "black", main = "B: Annual transition probability (Spline, 1 knot, normal scale)", 
         xlab = "time", ylab = "annual transition probability (smoothed)", 
         ylim = c(0, ceiling(km_tp$maxtp_smooth*10)/10), xlim = c(0, ceiling(km$maxtime*10)/10))
    lines(cbind(time_pred[-1], tp_gr_1[, 13]), col = "5", lty = 1, lwd = 2)
    if (ngroups > 1) {
      lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), col = "lightgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 8)
      lines(cbind(time_pred[-1], tp_gr_2[, 13]), col = "5", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
      lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), col = "darkgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 9)
      lines(cbind(time_pred[-1], tp_gr_3[, 13]), col = "5", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km_tp_gr_1$time, rev(km_tp_gr_1$time)), 
            c(km_tp_gr_1$tp_smooth_lower, rev(km_tp_gr_1$tp_smooth_upper)), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km_tp_gr_2$time, rev(km_tp_gr_2$time)), 
              c(km_tp_gr_2$tp_smooth_lower, rev(km_tp_gr_2$tp_smooth_upper)), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km_tp_gr_3$time, rev(km_tp_gr_3$time)), 
              c(km_tp_gr_3$tp_smooth_lower, rev(km_tp_gr_3$tp_smooth_upper)), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    abline(v = exp(spl_normal1$knots[2]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_normal1$knots[1]), lty = 3, lwd = 1)
    abline(v = exp(spl_normal1$knots[3]), lty = 3, lwd = 1)
    legend("topleft", legend = group_names, col = "5", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)

    # diagnostics
    plot(cbind(log(km$time), -qnorm(km$surv, mean = 0, sd = 1, lower.tail = TRUE, 
        log.p = FALSE)), pch = c(1, 8, 9)[km_names], col = c("black", "lightgrey", 
        "darkgrey")[km_names], cex = 0.6, main = "C: Diagnostic plot (Spline, 1 knot, normal scale)", 
        xlim = c(floor(min(log(km$time)[km$surv!=1])), ceiling(max(log(km$time)))),
        xlab = "LN(time)", ylab = "- standard normal quartiles")
    lines(cbind(log(time_pred), -qnorm(spl_normal1_pred[, 2], mean = 0, sd = 1, lower.tail = TRUE, 
        log.p = FALSE)), col = "5", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(log(time_pred), -qnorm(spl_normal1_pred[, 3], mean = 0, sd = 1, 
            lower.tail = TRUE, log.p = FALSE)), col = "5", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(log(time_pred), -qnorm(spl_normal1_pred[, 4], mean = 0, sd = 1, 
            lower.tail = TRUE, log.p = FALSE)), col = "5", lty = "5212", lwd = 2)
    }
    abline(v = spl_normal1$knots[2], lty = 2, lwd = 1.5)
    abline(v = spl_normal1$knots[1], lty = 3, lwd = 1)
    abline(v = spl_normal1$knots[3], lty = 3, lwd = 1)
    legend("topleft", legend = group_names, col = "5", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)
}, envir = PERSUADE)
```

##Spline normal 2 knots
```{r spline_norm2, echo=FALSE, out.height = "30%", fig.width=11, eval = PERSUADE$show_spline}
evalq({
    # 2 knots
    plot(km, lwd = 2, main = "A: Kaplan-Meier (Spline, 2 knots, normal scale)", xlab = "time", 
        ylab = "survival", col = c("black", "lightgrey", "darkgrey"))
    lines(cbind(time_pred, spl_normal2_pred[, 2]), col = "6", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(time_pred, spl_normal2_pred[, 3]), col = "6", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(time_pred, spl_normal2_pred[, 4]), col = "6", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km$time[km_names==1], rev(km$time[km_names==1])), 
            c(km$lower[km_names==1], rev(km$upper[km_names==1])), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km$time[km_names==2], rev(km$time[km_names==2])), 
              c(km$lower[km_names==2], rev(km$upper[km_names==2])), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km$time[km_names==3], rev(km$time[km_names==3])), 
              c(km$lower[km_names==3], rev(km$upper[km_names==3])), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    abline(v = exp(spl_normal2$knots[2]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_normal2$knots[3]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_normal2$knots[1]), lty = 3, lwd = 1)
    abline(v = exp(spl_normal2$knots[4]), lty = 3, lwd = 1)
    legend("bottomleft", legend = group_names, col = "6", lty = as.integer(c(1, "2222", 
                                                                             "5212")), cex = 0.8)

    # annual transition probabilities
    plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), lwd = 1, lty = 1, type = "b", 
         cex = 0.6, pch = 1, col = "black", main = "B: Annual transition probability (Spline, 2 knots, normal scale)", 
         xlab = "time", ylab = "annual transition probability (smoothed)", 
         ylim = c(0, ceiling(km_tp$maxtp_smooth*10)/10), xlim = c(0, ceiling(km$maxtime*10)/10))
    lines(cbind(time_pred[-1], tp_gr_1[, 14]), col = "6", lty = 1, lwd = 2)
    if (ngroups > 1) {
      lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), col = "lightgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 8)
      lines(cbind(time_pred[-1], tp_gr_2[, 14]), col = "6", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
      lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), col = "darkgrey", type = "b", 
            lwd = 1, cex = 0.6, pch = 9)
      lines(cbind(time_pred[-1], tp_gr_3[, 14]), col = "6", lty = "5212", lwd = 2)
    }
    # add shading confidence intervals
    polygon(c(km_tp_gr_1$time, rev(km_tp_gr_1$time)), 
            c(km_tp_gr_1$tp_smooth_lower, rev(km_tp_gr_1$tp_smooth_upper)), 
            col = adjustcolor("black", alpha.f = 0.3), border = NA)
    if (ngroups > 1) {
      polygon(c(km_tp_gr_2$time, rev(km_tp_gr_2$time)), 
              c(km_tp_gr_2$tp_smooth_lower, rev(km_tp_gr_2$tp_smooth_upper)), 
              col = adjustcolor("lightgrey", alpha.f = 0.3), border = NA)
    }
    if (ngroups > 2) {
      polygon(c(km_tp_gr_3$time, rev(km_tp_gr_3$time)), 
              c(km_tp_gr_3$tp_smooth_lower, rev(km_tp_gr_3$tp_smooth_upper)), 
              col = adjustcolor("darkgrey", alpha.f = 0.3), border = NA)
    }
    abline(v = exp(spl_normal2$knots[2]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_normal2$knots[3]), lty = 2, lwd = 1.5)
    abline(v = exp(spl_normal2$knots[1]), lty = 3, lwd = 1)
    abline(v = exp(spl_normal2$knots[4]), lty = 3, lwd = 1)
    legend("topleft", legend = group_names, col = "6", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)

    # diagnostics
    plot(cbind(log(km$time), -qnorm(km$surv, mean = 0, sd = 1, lower.tail = TRUE, 
        log.p = FALSE)), pch = c(1, 8, 9)[km_names], col = c("black", "lightgrey", 
        "darkgrey")[km_names], cex = 0.6, main = "C: Diagnostic plot (Spline, 2 knots, normal scale)",
        xlim = c(floor(min(log(km$time)[km$surv!=1])), ceiling(max(log(km$time)))),
        xlab = "LN(time)", ylab = "- standard normal quartiles")
    lines(cbind(log(time_pred), -qnorm(spl_normal2_pred[, 2], mean = 0, sd = 1, lower.tail = TRUE, 
        log.p = FALSE)), col = "6", lty = 1, lwd = 2)
    if (ngroups > 1) {
        lines(cbind(log(time_pred), -qnorm(spl_normal2_pred[, 3], mean = 0, sd = 1, 
            lower.tail = TRUE, log.p = FALSE)), col = "6", lty = "2222", lwd = 2)
    }
    if (ngroups > 2) {
        lines(cbind(log(time_pred), -qnorm(spl_normal2_pred[, 4], mean = 0, sd = 1, 
            lower.tail = TRUE, log.p = FALSE)), col = "6", lty = "5212", lwd = 2)
    }
    abline(v = spl_normal2$knots[2], lty = 2, lwd = 1.5)
    abline(v = spl_normal2$knots[3], lty = 2, lwd = 1.5)
    abline(v = spl_normal2$knots[1], lty = 3, lwd = 1)
    abline(v = spl_normal2$knots[4], lty = 3, lwd = 1)
    legend("topleft", legend = group_names, col = "6", lty = as.integer(c(1, "2222", 
                                                                          "5212")), cex = 0.8)
}, envir = PERSUADE)
```

\newpage

#Validity of long-term extrapolation?
What model(s) is/are more appropriate for long-term extrapolation?
Are/is the selected model(s) plausible in comparison with general population mortality?

##`r paste("Group", PERSUADE$group_names[1])`
```{r validate_extrapolation1, echo=FALSE, out.height = "29%", fig.width=11, comment=NA}
evalq({
    # parametric curves
    plot(km[1], lwd = 2, col = "black", main = "A: Kaplan-Meier (parametric curves)", 
        xlab = "time", ylab = "survival", xlim = c(0, time_horizon))
    lines(cbind(time_pred, expo_pred[, 2]), col = "1", lty = 1, lwd = 1)
    lines(cbind(time_pred, weib_pred[, 2]), col = "2", lty = 1, lwd = 1)
    lines(cbind(time_pred, gom_pred[, 2]), col = "3", lty = 1, lwd = 1)
    lines(cbind(time_pred, lnorm_pred[, 2]), col = "4", lty = 1, lwd = 1)
    lines(cbind(time_pred, llog_pred[, 2]), col = "5", lty = 1, lwd = 1)
    lines(cbind(time_pred, gam_pred[, 2]), col = "6", lty = 1, lwd = 1)
    lines(cbind(time_pred, ggam_pred[, 2]), col = "7", lty = 1, lwd = 1)
    legend("topright", legend = lbls, col = 1:7, lty = 1, cex = 0.8)
    # spline curves
    if (show_spline == TRUE) {
        plot(km[1], lwd = 2, col = "black", main = "B: Kaplan-Meier (spline curves)", 
            xlab = "time", ylab = "survival", xlim = c(0, time_horizon))
        lines(cbind(time_pred, spl_hazard1_pred[, 2]), col = "1", lty = 1, lwd = 1)
        lines(cbind(time_pred, spl_hazard2_pred[, 2]), col = "2", lty = 1, lwd = 1)
        lines(cbind(time_pred, spl_odds1_pred[, 2]), col = "3", lty = 1, lwd = 1)
        lines(cbind(time_pred, spl_odds2_pred[, 2]), col = "4", lty = 1, lwd = 1)
        lines(cbind(time_pred, spl_normal1_pred[, 2]), col = "5", lty = 1, lwd = 1)
        lines(cbind(time_pred, spl_normal2_pred[, 2]), col = "6", lty = 1, lwd = 1)
        legend("topright", legend = lbls_spline, col = 1:6, lty = 1, cex = 0.8)
    }
}, envir = PERSUADE)
kable_styling(kable(t(round(PERSUADE$extrapolation_gr_1[1 + PERSUADE$time_pred_surv_table, 
    ], 3))[-1, ], col.names = paste("T=", PERSUADE$extrapolation_gr_1[1 + PERSUADE$time_pred_surv_table, 
    ][, 1]), format = "latex", booktabs = TRUE, linesep = ""), latex_options = c("scale_down", 
    "striped"), position = "left")

evalq({
    # parametric curves
    plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), main = "C: Annual transition probability (parametric curves)", 
        xlab = "time", ylab = "annual transition probability", ylim = c(0, 1), xlim = c(0, time_horizon),
        col = "black", type = "l", lwd = 2)
    lines(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth_lower), col = "black", type = "l", 
        lwd = 2, lty = 2)
    lines(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth_upper), col = "black", type = "l", 
        lwd = 2, lty = 2)
    lines(cbind(time_pred[-1], tp_gr_1[, 2]), col = "1", lty = 1, lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_1[, 3]), col = "2", lty = 1, lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_1[, 4]), col = "3", lty = 1, lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_1[, 5]), col = "4", lty = 1, lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_1[, 6]), col = "5", lty = 1, lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_1[, 7]), col = "6", lty = 1, lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_1[, 8]), col = "7", lty = 1, lwd = 1)
    legend("topleft", legend = lbls, col = 1:7, lty = 1, cex = 0.8)
    # spline curves
    if (show_spline == TRUE) {
        plot(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth), main = "D: Annual transition probability (spline curves)", 
             xlab = "time", ylab = "annual transition probability", ylim = c(0, 1), xlim = c(0, time_horizon),
             col = "black", type = "l", lwd = 2)
        lines(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth_lower), col = "black", type = "l", 
            lwd = 2, lty = 2)
        lines(cbind(km_tp_gr_1$time, km_tp_gr_1$tp_smooth_upper), col = "black", type = "l", 
            lwd = 2, lty = 2)
        lines(cbind(time_pred[-1], tp_gr_1[, 9]), col = "1", lty = 1, lwd = 1)  
        lines(cbind(time_pred[-1], tp_gr_1[, 10]), col = "2", lty = 1, lwd = 1)
        lines(cbind(time_pred[-1], tp_gr_1[, 11]), col = "3", lty = 1, lwd = 1)
        lines(cbind(time_pred[-1], tp_gr_1[, 12]), col = "4", lty = 1, lwd = 1)
        lines(cbind(time_pred[-1], tp_gr_1[, 13]), col = "5", lty = 1, lwd = 1)
        lines(cbind(time_pred[-1], tp_gr_1[, 14]), col = "6", lty = 1, lwd = 1)
        legend("topleft", legend = lbls_spline, col = 1:6, lty = 1, cex = 0.8)
    }
}, envir = PERSUADE)
suppressWarnings(kable_styling(kable(descr(PERSUADE$tp_gr_1[, 2:PERSUADE$cols_extr], 
    stats = "fivenum", transpose = TRUE), format = "latex", booktabs = TRUE, linesep = ""), 
    latex_options = "striped", position = "left"))
```

\newpage

##`r paste("Group", PERSUADE$group_names[2])`
```{r validate_extrapolation2, echo=FALSE, out.height = "29%", fig.width=11, eval=if (PERSUADE$ngroups>1) {TRUE} else {FALSE}, comment=NA}
evalq({
    # parametric curves
    plot(km[2], lwd = 2, col = "black", main = "A: Kaplan-Meier (parametric curves)", 
        xlab = "time", ylab = "survival", xlim = c(0, time_horizon))
    lines(cbind(time_pred, expo_pred[, 3]), col = "1", lty = "2222", lwd = 1)
    lines(cbind(time_pred, weib_pred[, 3]), col = "2", lty = "2222", lwd = 1)
    lines(cbind(time_pred, gom_pred[, 3]), col = "3", lty = "2222", lwd = 1)
    lines(cbind(time_pred, lnorm_pred[, 3]), col = "4", lty = "2222", lwd = 1)
    lines(cbind(time_pred, llog_pred[, 3]), col = "5", lty = "2222", lwd = 1)
    lines(cbind(time_pred, gam_pred[, 3]), col = "6", lty = "2222", lwd = 1)
    lines(cbind(time_pred, ggam_pred[, 3]), col = "7", lty = "2222", lwd = 1)
    legend("topright", legend = lbls, col = 1:7, lty = "2222", cex = 0.8)
    # spline curves
    if (show_spline == TRUE) {
        plot(km[2], lwd = 2, col = "black", main = "B: Kaplan-Meier (spline curves)", 
            xlab = "time", ylab = "survival", xlim = c(0, time_horizon))
        lines(cbind(time_pred, spl_hazard1_pred[, 3]), col = "1", lty = "2222", lwd = 1)
        lines(cbind(time_pred, spl_hazard2_pred[, 3]), col = "2", lty = "2222", lwd = 1)
        lines(cbind(time_pred, spl_odds1_pred[, 3]), col = "3", lty = "2222", lwd = 1)
        lines(cbind(time_pred, spl_odds2_pred[, 3]), col = "4", lty = "2222", lwd = 1)
        lines(cbind(time_pred, spl_normal1_pred[, 3]), col = "5", lty = "2222", lwd = 1)
        lines(cbind(time_pred, spl_normal2_pred[, 3]), col = "6", lty = "2222", lwd = 1)
        legend("topright", legend = lbls_spline, col = 1:6, lty = "2222", cex = 0.8)
    }
}, envir = PERSUADE)
kable_styling(kable(t(round(PERSUADE$extrapolation_gr_2[1 + PERSUADE$time_pred_surv_table, 
    ], 3))[-1, ], col.names = paste("T=", PERSUADE$extrapolation_gr_2[1 + PERSUADE$time_pred_surv_table, 
    ][, 1]), format = "latex", booktabs = TRUE, linesep = ""), latex_options = c("scale_down", 
    "striped"), position = "left")

evalq({
    # parametric curves
    plot(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), main = "C: Annual transition probability (parametric curves)", 
        xlab = "time", ylab = "annual transition probability", ylim = c(0, 1), xlim = c(0, time_horizon),
        col = "black", type = "l", lwd = 2)
    lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth_lower), col = "black", type = "l", 
        lwd = 2, lty = 2)
    lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth_upper), col = "black", type = "l", 
        lwd = 2, lty = 2)
    lines(cbind(time_pred[-1], tp_gr_2[, 2]), col = "1", lty = "2222", lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_2[, 3]), col = "2", lty = "2222", lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_2[, 4]), col = "3", lty = "2222", lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_2[, 5]), col = "4", lty = "2222", lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_2[, 6]), col = "5", lty = "2222", lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_2[, 7]), col = "6", lty = "2222", lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_2[, 8]), col = "7", lty = "2222", lwd = 1)
    legend("topleft", legend = lbls, col = 1:7, lty = "2222", cex = 0.8)
    # spline curves
    if (show_spline == TRUE) {
        plot(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth), main = "D: Annual transition probability (spline curves)", 
             xlab = "time", ylab = "annual transition probability", ylim = c(0, 1), xlim = c(0, time_horizon),
             col = "black", type = "l", lwd = 2)
        lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth_lower), col = "black", type = "l", 
            lwd = 2, lty = 2)
        lines(cbind(km_tp_gr_2$time, km_tp_gr_2$tp_smooth_upper), col = "black", type = "l", 
            lwd = 2, lty = 2)
        lines(cbind(time_pred[-1], tp_gr_2[, 9]), col = "1", lty = "2222", lwd = 1)  
        lines(cbind(time_pred[-1], tp_gr_2[, 10]), col = "2", lty = "2222", lwd = 1)
        lines(cbind(time_pred[-1], tp_gr_2[, 11]), col = "3", lty = "2222", lwd = 1)
        lines(cbind(time_pred[-1], tp_gr_2[, 12]), col = "4", lty = "2222", lwd = 1)
        lines(cbind(time_pred[-1], tp_gr_2[, 13]), col = "5", lty = "2222", lwd = 1)
        lines(cbind(time_pred[-1], tp_gr_2[, 14]), col = "6", lty = "2222", lwd = 1)
        legend("topleft", legend = lbls_spline, col = 1:6, lty = "2222", cex = 0.8)
    }
}, envir = PERSUADE)
suppressWarnings(kable_styling(kable(descr(PERSUADE$tp_gr_2[, 2:PERSUADE$cols_extr], 
    stats = "fivenum", transpose = TRUE), format = "latex", booktabs = TRUE, linesep = ""), 
    latex_options = "striped", position = "left"))
```

\newpage

##`r paste("Group", PERSUADE$group_names[3])`
```{r validate_extrapolation3, echo=FALSE, out.height = "29%", fig.width=11, eval=if (PERSUADE$ngroups>2) {TRUE} else {FALSE}, comment=NA}
evalq({
    # parametric curves
    plot(km[3], lwd = 2, col = "black", main = "A: Kaplan-Meier (parametric curves)", 
        xlab = "time", ylab = "survival", xlim = c(0, time_horizon))
    lines(cbind(time_pred, expo_pred[, 4]), col = "1", lty = "5212", lwd = 1)
    lines(cbind(time_pred, weib_pred[, 4]), col = "2", lty = "5212", lwd = 1)
    lines(cbind(time_pred, gom_pred[, 4]), col = "3", lty = "5212", lwd = 1)
    lines(cbind(time_pred, lnorm_pred[, 4]), col = "4", lty = "5212", lwd = 1)
    lines(cbind(time_pred, llog_pred[, 4]), col = "5", lty = "5212", lwd = 1)
    lines(cbind(time_pred, gam_pred[, 4]), col = "6", lty = "5212", lwd = 1)
    lines(cbind(time_pred, ggam_pred[, 4]), col = "7", lty = "5212", lwd = 1)
    legend("topright", legend = lbls, col = 1:7, lty = "5212", cex = 0.8)
    # spline curves
    if (show_spline == TRUE) {
        plot(km[3], lwd = 2, col = "black", main = "B: Kaplan-Meier (spline curves)", 
            xlab = "time", ylab = "survival", xlim = c(0, time_horizon))
        lines(cbind(time_pred, spl_hazard1_pred[, 4]), col = "1", lty = "5212", lwd = 1)
        lines(cbind(time_pred, spl_hazard2_pred[, 4]), col = "2", lty = "5212", lwd = 1)
        lines(cbind(time_pred, spl_odds1_pred[, 4]), col = "3", lty = "5212", lwd = 1)
        lines(cbind(time_pred, spl_odds2_pred[, 4]), col = "4", lty = "5212", lwd = 1)
        lines(cbind(time_pred, spl_normal1_pred[, 4]), col = "5", lty = "5212", lwd = 1)
        lines(cbind(time_pred, spl_normal2_pred[, 4]), col = "6", lty = "5212", lwd = 1)
        legend("topright", legend = lbls_spline, col = 1:6, lty = "5212", cex = 0.8)
    }
}, envir = PERSUADE)
kable_styling(kable(t(round(PERSUADE$extrapolation_gr_3[1 + PERSUADE$time_pred_surv_table, 
    ], 3))[-1, ], col.names = paste("T=", PERSUADE$extrapolation_gr_3[1 + PERSUADE$time_pred_surv_table, 
    ][, 1]), format = "latex", booktabs = TRUE, linesep = ""), latex_options = c("scale_down", 
    "striped"), position = "left")

evalq({
    # parametric curves
    plot(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), main = "C: Annual transition probability (parametric curves)", 
        xlab = "time", ylab = "annual transition probability", ylim = c(0, 1), xlim = c(0, time_horizon),
        col = "black", type = "l", lwd = 2)
    lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth_lower), col = "black", type = "l", 
        lwd = 2, lty = 2)
    lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth_upper), col = "black", type = "l", 
        lwd = 2, lty = 2)
    lines(cbind(time_pred[-1], tp_gr_3[, 2]), col = "1", lty = "5212", lwd = 1)  
    lines(cbind(time_pred[-1], tp_gr_3[, 3]), col = "2", lty = "5212", lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_3[, 4]), col = "3", lty = "5212", lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_3[, 5]), col = "4", lty = "5212", lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_3[, 6]), col = "5", lty = "5212", lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_3[, 7]), col = "6", lty = "5212", lwd = 1)
    lines(cbind(time_pred[-1], tp_gr_3[, 8]), col = "7", lty = "5212", lwd = 1)
    legend("topleft", legend = lbls, col = 1:7, lty = "5212", cex = 0.8)
    # spline curves
    if (show_spline == TRUE) {
        plot(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth), main = "D: Annual transition probability (spline curves)", 
             xlab = "time", ylab = "annual transition probability", ylim = c(0, 1), xlim = c(0, time_horizon),
             col = "black", type = "l", lwd = 2)
        lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth_lower), col = "black", type = "l", 
            lwd = 2, lty = 2)
        lines(cbind(km_tp_gr_3$time, km_tp_gr_3$tp_smooth_upper), col = "black", type = "l", 
            lwd = 2, lty = 2)
        lines(cbind(time_pred[-1], tp_gr_3[, 9]), col = "1", lty = "5212", lwd = 1)  
        lines(cbind(time_pred[-1], tp_gr_3[, 10]), col = "2", lty = "5212", lwd = 1)
        lines(cbind(time_pred[-1], tp_gr_3[, 11]), col = "3", lty = "5212", lwd = 1)
        lines(cbind(time_pred[-1], tp_gr_3[, 12]), col = "4", lty = "5212", lwd = 1)
        lines(cbind(time_pred[-1], tp_gr_3[, 13]), col = "5", lty = "5212", lwd = 1)
        lines(cbind(time_pred[-1], tp_gr_3[, 14]), col = "6", lty = "5212", lwd = 1)
        legend("topleft", legend = lbls_spline, col = 1:6, lty = "5212", cex = 0.8)
    }
}, envir = PERSUADE)
suppressWarnings(kable_styling(kable(descr(PERSUADE$tp_gr_3[, 2:PERSUADE$cols_extr], 
    stats = "fivenum", transpose = TRUE), format = "latex", booktabs = TRUE, linesep = ""), 
    latex_options = "striped", position = "left"))
```

\newpage

#Session information
```{r si, echo=FALSE}
print(sessionInfo(), RNG = TRUE, locale = FALSE)
```
